### 2018-06-05 IEEEtran/article.tex
spelling corrections
finished an iteration of proofreading

% As a general rule, do not put math, special symbols or citations
% in the abstract
\begin{abstract}
Representing UML diagrams in logic can be of interest to a designer since the designer can easily miss errors or inefficiencies in the design if the diagrams grow sufficiently elaborate or numerous. In this article, we show a method to translate class diagrams and a corresponding set of sequence diagrams that model the behavior of a desired software system to FO($\cdot$), an extension of first-order predicate logic with inductive definitions, partial functions, aggregates and types. We show how the output theory may be used to verify consistency of a class diagram and detect the presence of certain design flaws in the class diagram. In addition, we show that the theory may be used to simulate system behavior as modelled in the sequence diagrams and how to verify that requirements regarding the output of a diagram are satisfied. We also evaluate performance in terms of execution time and the size of the grounding for each of these tasks.
\end{abstract}

@@

\section{Introduction}
% no \IEEEPARstart
UML\cite{RumbaughJames2005Tuml} is a modelling language used in software engineering to graphically model a design for a software system. Popular types of diagrams are the class diagram and the sequence diagram. However, when a class diagram grows too elaborate or the sequence diagrams too numerous, the designer can easily miss any design errors or flaws or unintended behavior. Since errors made in the design phase can be costly to correct if found late in the software production process, automated verification of consistency and checking for design flaws might help the designer spot issues they might otherwise miss.

@@

We will use a translation of a class diagram to verify whether the diagram is consistent, i.e. whether the FO($\cdot$) theory corresponding to the diagram has a model. We will also use it to detect certain types of design flaws which make it more difficult to gain a clear understanding of a class diagram.

@@

In \cite{BerardiDaniela2005RoUc}, the authors first provide a method to translate class diagrams to FO-theories. The vocabulary first defines: Predicates that express membership of a class; predicates that relate instances of a class to their respective values for class attributes; predicates that model operations on a class by relating the callee object, values for parameters and the result value corresponding to that callee object and those parameters values; and predicates that relate objects of a class to objects of another class according to associations defined in the class diagram. The rules contained in the output theory make use of the membership predicates to enforce correct typing for the other kinds of predicates and ensure that the predicates for attributes, operations and associations obey the multiplicites imposed on them by the diagrams. Finally, the theory models inheritance by introducing rules that state that all members of a subclass must also be members of the superclass. Note that these inheritance rules do not prevent members of one class being members of another class even if the diagram defines no inheritance relationship between those two classes, regardless of whether the developer wants to allow for that possibility. The rest of the paper describes a method to translate class diagrams to several kinds of description logics and how to use those translations to verify consistency of the diagram and to detect implicit properties which unintentionally hold in the diagram. In this article, we adapt the method for translating class diagrams to FO introduced by the authors to make use of logical types, thereby removing the need to introduce rules in the output theory that enforce the correct typing and to model inheritance.

@@

\begin{align*}
	&\forall{c_1}[Class_1]\cdots{}\forall{c_m}[Class_m](lowerBound_l \leq \\ &\#\{o_1[Class_l] : ClassOneand\cdots{}andClassM\\&(c_1, \cdots, o_l, \cdots, c_m)\} \leq upperBound_l).
\end{align*}

@@

To model inheritance, the vocabulary must simply declare that logical types corresponding to a subclass are subtypes of the logical type corresponding to the superclass. This means that every input structure for the output theory must include every object in the interpretation of a subtype in the interpretation of all the corresponding supertypes in order to be valid input for model expansion. It is possible to spare the user this inconvenience by abandoning the principle of one logical type per class and instead introducing a general logical type \textit{Object} and introducing another logical type \textit{ClassObject} which has one object for every class. We would then add a binary predicate \textit{StaticClass(ClassObject, Object)} in order to express membership of a class. Inductive definitions would then use the inheritance relationships contained in the diagram to calculate the correct class memberships for every object. However, this approach would require rewriting every rule introduced earlier in this section to be longer by enforcing correct typing, which would make the output theory less easily understandable. There would also have to be additional rules to enforce the correct type for every parameter for every attribute and association predicate. For these reasons, we chose not to implement this approach.

@@

\textit{BiAssocLow(ClassObject, ClassObject, ClassObject, nat)} indicates that the role corresponding to the class named in the third parameter has the lower bound specified in the fourth argument.

Similarly, \textit{BiAssocHigh(ClassObject, ClassObject, ClassObject, nat)} specifies the upper bound for the role corresponding to the third parameter.

@@

\section{Translating sequence diagrams to FO($\cdot$)}\label{sec:seq}
We will use the framework of linear time calculus\cite{BogaertsBart2014Sdsu} (LTC). LTC can model dynamic systems that change over time. Across each time step, only the inertial predicates that need be changed do so, while all other inertial predicates remain the same. If we model diagram variables, an instruction pointer, a stack level counter and a return point stack as inertial predicates, we have exactly the elements we need to translate mutually recursive sequence diagrams to FO($\cdot$) theories.

@@

\begin{itemize}
	\item \textit{SDPoint}: This serves as an instruction pointer. It is a constructed type that consists of labels for all instructions across all given sequence diagrams, landing points after each call to a sequence diagram, and hidden return instruction for sequence diagrams that model methods that return \textit{void}. \textit{SDPoint}s are named according to the pattern \textit{$diagramname\_\langle{}sequencenumber\rangle$}, except landing points after diagram calls, which are named according to \textit{$diagramname\_\langle{}sequencenumber\rangle$post}
	\item \textit{NextSD(SDPoint) : SDPoint}: A partial function which specifies which \textit{SDPoint} follows the given \textit{SDPoint}. All \textit{SDPoint}s except the last instruction for each diagram are members of the domain.
	\item \textit{SDPointAt(Time, SDPoint)}: The \textit{SDPoint} at any given time step.
	\item \textit{StackLevel}: A non-zero natural number that indicates the stack depth.
	\item \textit{CurrentStackLevel(Time) : StackLevel}: A total function which specifies what the stack depth is at any given time step.
	\item \textit{ReturnPoint(Time, StackLevel, SDPoint)}: This predicate contains the return instruction at the specified stack depth. If the execution reaches the final instruction of a diagram, the instruction counter should be set to the corresponding \textit{SDPoint} at the following time step.
	\item Predicates for every diagram variable named according to the pattern \textit{VariableNameT(Time, StackLevel, VariableType)}. These diagram variables include variables defined in an instruction, lifeline objects and arguments of calls.
\end{itemize}

@@

With these logical symbols, predicates and functions at our disposal, we are now equipped to translate sequence diagrams to FO($\cdot$) theories. Consider figure \ref{fig:recursion}. Diagram \ref{fig:methodOne} calls diagram \ref{fig:methodTwo}, whereupon diagram \ref{fig:methodTwo} calls itself two times and then returns. In this case, the members of \textit{SDPoint} are \textit{methodOne\_1}, \textit{methodOne\_2}, \textit{methodOne\_3}, \textit{methodOne\_3post}, \textit{methodOne\_4}, the hidden return instruction \textit{methodOne\_5}, \textit{methodTwo\_1}, \textit{methodTwo\_2}, \textit{methodTwo\_3}, \textit{methodTwo\_3post}, \textit{methodTwo\_4} and hidden return instruction \textit{methodTwo\_5}. We introduce the inertial predicates \textit{ObjT/3}, \textit{Obj2T/3}, \textit{FinishedT/3} and \textit{MTwoArgT/3}. We add causation rules corresponding to each instruction that contains an assignment to a variable or an object call. In the latter case, we add causation rules that set the correct callee object, call arguments, ensure that the value for \textit{SDPointAt/2} is the first instruction of the called diagram and set the return instruction for the next stack level to the hidden landing point after the call instruction. We also add causation rules corresponding to the final instruction of each diagram where the return instruction for the current stack depth is retrieved and set to be the value of \textit{SDPointAt/2} at the next time step. If any diagram modelled a non-void method, then we would add a causation rule corresponding to the landing point after a call of that diagram where the returned value is retrieved and is set to be the value of the assigned variable.

@@

The general approach with regard to parsing combined fragment then goes as follows. If the message immediately preceding the combined fragment currently under consideration is not part of a combined fragment, then determine which are the first possible messages of this fragment and note that the execution jumps from that preceding message to those first messages if their corresponding conditions hold. If the immediately preceding message is part of a combined fragment, then determine the possible exit instructions for that combined fragment and use that information to determine which final messages for the preceding fragment jump to which possible first messages for the current fragment under which conditions. Finally, check if the current fragment is nested in one or more loop combined fragments and determine what instructions the execution might jump to when re-entering the loops. The output of this procedure is a set of causation rules for \textit{SDPointAt/3}.

@@

We simulated a game with two heaps. One of the heaps contains two objects and the other contains three objects. The first player removes all three objects from the second heap. The second player then removes one object from the first heap. Finally, the first player has no choice but to remove the last object and loses.
Figure \ref{fig:grounding} contains a graph of the size of the grounding reported by IDP per step. Figure \ref{fig:ms} shows a graph of the cumulative execution time per step measured in milliseconds. We can see that the grounding size was either one or three for most steps, with six exceptions. The execution time graph shows a linear curve, indicating that the time interval between steps remained stable. It also shows that the total duration of the game was approximately 18 minutes, which is fairly long for a simple game like Nim.

@@

Each run did indeed lead to the expected answer. For each diagram, the runs were similar in terms of execution time and grounding size. For \textit{isEmpty()}, the execution time is approximately 21.5 seconds and the size of the grounding is 578,306. For \textit{allHeapsEmpty()} given two heaps, the execution time is approximately 24 seconds and the grounding size is 2,249,098.

@@

\begin{figure*}[!t]
	\centering
	\subfloat[Sequence diagram from \textit{takeTurn()}]{\includegraphics[width=0.5\textwidth]{takeTurn}%
		\label{fig:takeTurn}}
	\hfil
	\subfloat[Sequence diagram for \textit{take(int)}]{\includegraphics[width=0.5\textwidth]{take}}%
	\label{fig:take}
	\caption{Sequence diagrams for \textit{allHeapsEmpty()} and \textit{isEmpty()}}
	\label{fig:tt-take}
\end{figure*}

>>>>

% As a general rule, do not put math, special symbols or citations
% in the abstract
\begin{abstract}
Representing UML diagrams in logic can benefit a designer since the designer can easily miss errors or inefficiencies in the design if the diagrams grow sufficiently elaborate or numerous. In this article, we show a method to translate class diagrams and a corresponding set of sequence diagrams that model the behavior of a desired software system to FO($\cdot$), an extension of first-order predicate logic with inductive definitions, partial functions, aggregates and types. We show how the output theory may be used to verify consistency of a class diagram and detect the presence of certain design flaws in the class diagram. In addition, we show that the theory may be used to simulate system behavior as modelled in the sequence diagrams and how to verify that requirements regarding the output of a diagram are satisfied. We also evaluate performance in terms of execution time and the size of the grounding for each of these tasks.
\end{abstract}

@@

\section{Introduction}
% no \IEEEPARstart
UML\cite{RumbaughJames2005Tuml} is a modelling language used in software engineering to graphically model a design for a software system. Popular types of diagrams are the class diagram and the sequence diagram. However, when a class diagram grows too elaborate or the sequence diagrams too numerous, the designer can easily miss any design errors or flaws or unintended behavior. Since errors made in the design phase can be costly to correct if found late in the software production process, automated verification of consistency and detection of the presence of design flaws might help the designer spot issues they might otherwise miss.

@@

We will use a translation of a class diagram to verify whether the diagram is consistent, i.e. whether the FO($\cdot$) theory corresponding to the diagram has a model. We will also use an alternative translation to detect certain types of design flaws which make it more difficult to gain a clear understanding of a class diagram.

@@

In \cite{BerardiDaniela2005RoUc}, the authors first provide a method to translate class diagrams to FO-theories. The vocabulary first defines: Predicates that express membership of a class; predicates that relate instances of a class to their respective values for class attributes; predicates that model operations on a class by relating the callee object, values for parameters and the result value corresponding to that callee object and those parameters values; and predicates that relate objects of a class to objects of another class according to associations defined in the class diagram. The rules contained in the output theory make use of the membership predicates to enforce correct typing for the other kinds of predicates and ensure that the predicates for attributes, operations and associations obey the multiplicites imposed on them by the diagrams. Finally, the theory models inheritance by introducing rules that state that all members of a subclass must also be members of the superclass. Note that these inheritance rules do not prevent members of one class being members of another class even if the diagram defines no inheritance relationship between those two classes, regardless of whether the developer wants to allow for that possibility. The rest of the paper describes a method to translate class diagrams to several kinds of description logics and how to use those translations to verify consistency of the diagram and to detect implicit properties which unintentionally hold in the diagram. In this article, we adapt the method for translating class diagrams to FO introduced by the authors to make use of logical types, thereby removing the need to introduce rules in the output theory that enforce the correct typing and rules that model inheritance.

@@

\begin{align*}
	&\forall{c_1}[Class_1]\cdots{}\forall{c_m}[Class_m](lowerBound_l \leq \\ &\#\{o_l[Class_l] : ClassOneand\cdots{}andClassM\\&(c_1, \cdots, o_l, \cdots, c_m)\} \leq upperBound_l).
\end{align*}

@@

To model inheritance, the vocabulary must simply declare that logical types corresponding to a subclass are subtypes of the logical type corresponding to the superclass. This means that every input structure for the output theory must include all members of the interpretation of a subtype in the interpretation of all the corresponding supertypes in order to be valid input for model expansion. It is possible to spare the user this inconvenience by abandoning the principle of one logical type per class and instead introducing a general logical type \textit{Object} and introducing another logical type \textit{ClassObject} which has one object for every class. We would then add a binary predicate \textit{StaticClass(ClassObject, Object)} in order to express membership of a class. Inductive definitions would then use the inheritance relationships contained in the diagram to calculate the correct class memberships for every object. However, this approach would require rewriting every rule introduced earlier in this section to be longer by enforcing correct typing, which would make the output theory more difficult to understand. There would also have to be additional rules to enforce the correct type for every argument of every attribute and association predicate. For these reasons, we chose not to implement this approach.

@@

\textit{BiAssocLow(ClassObject, ClassObject, ClassObject, nat)} indicates that the role corresponding to the class named in the third argument has the lower bound specified in the fourth argument.

Similarly, \textit{BiAssocHigh(ClassObject, ClassObject, ClassObject, nat)} specifies the upper bound for the role corresponding to the argument parameter.

@@

\section{Translating sequence diagrams to FO($\cdot$)}\label{sec:seq}
We will use the framework of linear time calculus\cite{BogaertsBart2014Sdsu} (LTC). LTC can model dynamic systems that change over time. Across each time step, only the inertial predicates that need be changed to the diagram do so, while all other inertial predicates remain the same. If we model diagram variables, an instruction pointer, a stack level counter and a return point stack as inertial predicates, we have exactly the elements we need to translate mutually recursive sequence diagrams to FO($\cdot$) theories.

@@

\begin{itemize}
	\item \textit{SDPoint}: This serves as an instruction pointer. It is a constructed type that consists of labels for all instructions across all given sequence diagrams, landing points after each call to a sequence diagram, and additional hidden last instructions we add for sequence diagrams that model methods that return \textit{void}. \textit{SDPoint}s are named according to the pattern \textit{$diagramname\_\langle{}sequencenumber\rangle$}, except landing points after diagram calls, which are named according to \textit{$diagramname\_\langle{}sequencenumber\rangle$post}
	\item \textit{NextSD(SDPoint) : SDPoint}: A partial function that specifies which \textit{SDPoint} follows the given \textit{SDPoint}. All \textit{SDPoint}s except the last instruction for each diagram are members of the domain.
	\item \textit{SDPointAt(Time, SDPoint)}: The \textit{SDPoint} at any given time step.
	\item \textit{StackLevel}: A non-zero natural number that indicates the stack depth.
	\item \textit{CurrentStackLevel(Time) : StackLevel}: A total function which specifies what the stack depth is at any given time step.
	\item \textit{ReturnPoint(Time, StackLevel, SDPoint)}: This predicate contains the return instruction at the specified stack depth. If the execution reaches the final instruction of a diagram, the instruction counter should be set to the corresponding \textit{SDPoint} at the following time step.
	\item Predicates for every diagram variable named according to the pattern \textit{VariableNameT(Time, StackLevel, VariableType)}. These diagram variables include variables defined in an instruction, lifeline objects and arguments of calls.
\end{itemize}

@@

With these logical symbols, predicates and functions at our disposal, we are now equipped to translate sequence diagrams to FO($\cdot$) theories. Consider figure \ref{fig:recursion}. Diagram \ref{fig:methodOne} calls diagram \ref{fig:methodTwo}, whereupon diagram \ref{fig:methodTwo} calls itself two times and then returns. In this case, the members of \textit{SDPoint} are \textit{methodOne\_1}, \textit{methodOne\_2}, \textit{methodOne\_3}, \textit{methodOne\_3post}, \textit{methodOne\_4}, the hidden final instruction \textit{methodOne\_5}, \textit{methodTwo\_1}, \textit{methodTwo\_2}, \textit{methodTwo\_3}, \textit{methodTwo\_3post}, \textit{methodTwo\_4} and hidden final instruction \textit{methodTwo\_5}. We introduce the inertial predicates \textit{ObjT/3}, \textit{Obj2T/3}, \textit{FinishedT/3} and \textit{MTwoArgT/3}. We add causation rules corresponding to each instruction that contains an assignment to a variable or an object call. In the latter case, we add causation rules that set the correct callee object, call arguments, ensure that the value for \textit{SDPointAt/2} is the first instruction of the called diagram and set the return instruction for the next stack level to the hidden landing point after the call instruction. We also add causation rules corresponding to the final instruction of each diagram where the return instruction for the current stack depth is retrieved and set to be the value of \textit{SDPointAt/2} at the next time step. If any diagram modelled a non-void method, then we would add a causation rule corresponding to the landing point after a call of that diagram where the returned value is retrieved and is set to be the value of the assigned variable.

@@

The general approach with regard to parsing combined fragment then goes as follows. If the message immediately preceding the combined fragment currently under consideration is not part of a combined fragment, then determine which are the first possible messages of this fragment and note that the execution jumps from that preceding message to those first messages if their corresponding conditions hold. If the immediately preceding message is part of a combined fragment, then determine the possible exit instructions for that combined fragment and use that information to determine which final messages for the preceding fragment jump to which possible first messages for the current fragment under which conditions. Finally, check if the current fragment is a loop combined fragment and/or is nested in one or more loop combined fragments and determine what instructions the execution might jump to when re-entering the loops. The output of this procedure is a set of causation rules for \textit{SDPointAt/3}.

@@

We simulated a game with two heaps. One of the heaps contains two objects and the other contains three objects. The first player removes all three objects from the second heap. The second player then removes one object from the first heap. Finally, the first player has no choice but to remove the last object and loses.
Figure \ref{fig:grounding} contains a graph of the size of the grounding reported by IDP per step. Figure \ref{fig:ms} shows a graph of the cumulative execution time per step measured in milliseconds. We can see that the grounding size was either one or three for most steps, with six exceptions. However, IDP was using a total of approximately 4.84 gigabyte of virtual memory at the final time step, which was also the greatest usage during this simulation run. The execution time graph shows a linear curve, indicating that the time interval between steps remained stable. It also shows that the total duration of the game was approximately 18 minutes, which is fairly long for a simple game like Nim.

@@

The input structure for \textit{isEmpty()} specifies that there are seven time steps in total, while for \textit{allHeapsEmpty()}, there are 22 time steps in total, which suffices to reach a conclusive answer for two heaps. Each run did indeed lead to the expected answer. For each diagram, the runs were similar in terms of execution time and grounding size. For \textit{isEmpty()}, the execution time is approximately 29 seconds and the size of the grounding is 578,306. For \textit{allHeapsEmpty()} given two heaps, the execution time is approximately 1 minute and 28 seconds and the grounding size is 2,249,098.

@@

\begin{figure*}[!t]
	\centering
	\subfloat[Sequence diagram from \textit{takeTurn()}]{\includegraphics[width=0.5\textwidth]{takeTurn}%
		\label{fig:takeTurn}}
	\hfil
	\subfloat[Sequence diagram for \textit{take(int)}]{\includegraphics[width=0.5\textwidth]{take}}%
	\label{fig:take}
	\caption{Sequence diagrams for \textit{takeTurn()} and \textit{take(int)}}
	\label{fig:tt-take}
\end{figure*}

### 2018-06-05 chap-consistentie/chap-consistentie.tex
finished an iteration of proofreading

\section{Voorstellen van klassehi\"erarchi\"een}\label{sec:hierarchies}
Aangezien FO($\cdot$) een getypeerde logica is, is het eenvoudig om klassehi\"erarchie\"een voor te stellen. Als volgens het diagram klasse \textit{X} een subklasse is van klasse \textit{Y}, dan is het voldoende om in het uitvoervocabularium te noteren dat logisch type \textit{Y} een supertype is van logisch type \textit{X}. Het is mogelijk om op die manier een ketting van overervingsrelaties te modelleren en zo een hi\"erarchie te verkrijgen. In IDP in het bijzonder is het mogelijk dat een logisch type een subtype is van meer dan \'e\'en type, dus kan men ook meervoudige overerving modelleren.

Dit heeft echter een gevolg voor interpretaties van het logisch type overeenkomstig met een superklasse in een mogelijke structuur volgens het uitvoervocabularium. Indien klasse \textit{Z} een subklasse is van klasse \textit{Y}, dan moeten alle objecten die behoren tot klasse \textit{Z} ook behoren tot klasse \textit{X}, en zo ook moeten alle objecten die behoren tot klasses \textit{Z} en \textit{X} behoren tot klasse \textit{Y}.

@@

Er zijn ook twee predicaten nodig om overervingsrelaties tussen klasses te modelleren: \textit{IsDirectSupertypeOf(ClassObject, ClassObject)} om directe overervingsrelaties te capteren; en \textit{IsSupertypeOf(ClassObject, ClassObject)} dat de transitieve sluiting voor \textit{IsDirectSupertypeOf} voorstelt. Aangezien het in predicatenlogica onmogelijk is om een algemene voorstelling voor transitieve sluitingen uit te drukken, maken we gebruik van inductieve definities\cite{DeCatBroes2014PLaa}:

@@

We hebben echter niet voor deze aanpak gekozen voor twee redenen.
Attribuutpredicaten, operatiepredicaten en associatiepredicaten zouden logisch type \textit{Object} als argumenten moeten hebben. Dit betekent dat er nieuwe regels nodig zijn die afdwingen dat die objecten van het juiste type zijn. Grotere theorie\"en leiden tot meer rekentijd en geheugengebruik.
De multipliciteitsregels zouden ook herschreven moeten worden om gebruik te maken van \textit{StaticClass}. Dit maakt zinnen over de hele lijn langer en zorgt ervoor dat ze moeilijker te begrijpen zijn.

>>>>

\section{Voorstellen van klassehi\"erarchi\"een}\label{sec:hierarchies}
Aangezien FO($\cdot$) een getypeerde logica is, is het eenvoudig om klassehi\"erarchie\"een voor te stellen. Als volgens het diagram klasse \textit{X} een subklasse is van klasse \textit{Y}, dan is het voldoende om in het uitvoervocabularium te noteren dat logisch type \textit{Y} een supertype is van logisch type \textit{X}. Het is mogelijk om op die manier een keten van overervingsrelaties te modelleren en zo een hi\"erarchie te verkrijgen. In IDP in het bijzonder is het mogelijk dat een logisch type een subtype is van meer dan \'e\'en type, dus kan men ook meervoudige overerving modelleren.

Dit heeft echter een gevolg voor interpretaties van het logisch type overeenkomstig met een superklasse in een mogelijke structuur volgens het uitvoervocabularium. Indien klasse \textit{Z} een subklasse is van klasse \textit{X} uit de vorige paragraaf, dan moeten alle objecten die behoren tot klasse \textit{Z} ook behoren tot klasse \textit{X}, en zo ook moeten alle objecten die behoren tot klasses \textit{Z} en \textit{X} behoren tot klasse \textit{Y}.

@@

Er zijn ook twee predicaten nodig om overervingsrelaties tussen klasses te modelleren: \textit{IsDirectSupertypeOf(ClassObject, ClassObject)} om directe overervingsrelaties voor te stellen; en \textit{IsSupertypeOf(ClassObject, ClassObject)} dat de transitieve sluiting voor \textit{IsDirectSupertypeOf} voorstelt. Aangezien het in predicatenlogica onmogelijk is om een algemene voorstelling voor transitieve sluitingen uit te drukken, maken we gebruik van inductieve definities\cite{DeCatBroes2014PLaa}:

@@

We hebben echter niet voor deze aanpak gekozen voor twee redenen.
Attribuutpredicaten, operatiepredicaten en associatiepredicaten zouden logisch type \textit{Object} als argumenten moeten hebben. Dit betekent dat er nieuwe regels nodig zijn die afdwingen dat die objecten van het juiste type zijn. Grotere theorie\"en leiden tot een grotere rekentijd en geheugengebruik bij de uitvoering van redeneertaken.
De multipliciteitsregels zouden ook herschreven moeten worden om gebruik te maken van \textit{StaticClass}. Dit maakt zinnen over de hele lijn langer en zorgt ervoor dat ze moeilijker te begrijpen zijn.

### 2018-06-05 chap-evaluatie/chap-evaluatie.tex
finished an iteration of proofreading

De klasse \textit{Game} stelt het concept van een spel voor. \textit{p1Win} geeft aan of speler 1 heeft gewonnen en \textit{gameFinished} geeft aan of het spel volledig gespeeld is. \textit{allHeapsEmpty} vraagt op of alle \textit{Heap}s leeg zijn. \textit{takeTurn} doet een speler zijn beurt spelen. \textit{play(boolean)} is de methode die het verloop van het spel regelt.

De klasse \textit{Heap} stelt stapels van objecten voor. \textit{numObjects} is het aantal objecten dat een stapel bevat. \textit{isEmpty} geeft aan of de stapel leeg is en \textit{take(int)} neemt een aantal objecten weg van een stapel.

@@

Uit deze resultaten blijkt dat we met de uitvoertheorie het spel kunnen spelen, maar het tijdsinterval tussen beurten lang is. De grootte van de \textit{grounding} blijft beperkt.

\section{Verificatie van de correctheid van \textit{isEmpty}}

We willen controleren of het diagram in figuur \ref{fig:nim-isEmpty} een positief antwoord geeft als en slechts als de gegeven stapel leeg is. We werken met een invoerstructuur die stelt dat de uitvoering begint met de eerste instructie van \textit{isEmpty} en eindigt na de laatste instructie van hetzelfde diagram. Om de correctheid te controleren we vier zinnen op:

@@

Er is een gelijkaardig tijdsverloop en \textit{grounding}-grootte voor alle vier de uitvoeringen. Het duurt ongeveer 21,5 seconden en de \textit{grounding}-grootte is 578 306.

We merken op dat het relatief\todo{tegenover?} lang duurt om modelexpansie uit te voeren voor dit diagram. De \textit{grounding} is ook omvangrijk.

\section{Verificatie van de correctheid van \textit{allHeapsEmpty}}

We willen controleren of het diagram in figuur \ref{fig:nim-allHeapsEmpty} een positief antwoord geeft als alle stapels leeg zijn en een negatief antwoord als minstens \'e\'en stapel niet leeg is. We werken met een invoerstructuur die aangeeft dat de uitvoering begint met de eerste instructie van \textit{allHeapsEmpty} en eindigt na de laatste instructie van dat diagram. We stellen vier zinnen op:

@@

De uitvoeringstijd en \textit{grounding}-grootte zijn gelijkaardig voor alle vier uitvoeringen. De uitvoering duurt ongeveer 24 seconden en de \textit{grounding}-grootte is 2 249 098.

@@

Het is mogelijk om voor een diagram bepaalde gewenste eigenschappen te controleren. Dit gaat echter gepaard met grote \textit{groundings}.

>>>>

De klasse \textit{Game} stelt het concept van een spel voor. \textit{p1Win} geeft aan of speler 1 heeft gewonnen en \textit{gameFinished} geeft aan of het spel volledig gespeeld is. \textit{allHeapsEmpty()} vraagt op of alle \textit{Heap}s leeg zijn. \textit{takeTurn()} doet een speler zijn beurt spelen. \textit{play(boolean)} is de methode die het verloop van het spel regelt.

De klasse \textit{Heap} stelt stapels van objecten voor. \textit{numObjects} is het aantal objecten dat een stapel bevat. \textit{isEmpty()} geeft aan of de stapel leeg is en \textit{take(int)} neemt een aantal objecten weg van een stapel.

@@

Uit deze resultaten blijkt dat we met de uitvoertheorie het spel kunnen spelen, maar het tijdsinterval tussen beurten groot is. De grootte van de \textit{grounding} blijft beperkt. We merkten echter wel dat IDP na de laatste stap ongeveer 4,84 gigabyte in totaal aan virtueel geheugen in beslag nam, wat ook het grootste meetresultaat voor geheugengebruik over de hele simulatie heen was.

\section{Verificatie van de correctheid van \textit{isEmpty()}}

We willen controleren of het diagram in figuur \ref{fig:nim-isEmpty} een positief antwoord geeft als en slechts als de gegeven stapel leeg is. We werken met een invoerstructuur die stelt dat de uitvoering begint met de eerste instructie van \textit{isEmpty} en eindigt na de laatste instructie van hetzelfde diagram. De invoerstructuur specificeert dat er in totaal zeven tijdstappen zijn.

Om de correctheid te controleren we vier zinnen op:

@@

Er is een gelijkaardig tijdsverloop en \textit{grounding}-grootte voor alle vier de uitvoeringen. Het duurt ongeveer 29 seconden en de \textit{grounding}-grootte is 578 306.

We merken op dat het relatief lang duurt om modelexpansie uit te voeren voor dit diagram. De \textit{grounding} is ook omvangrijk.

\section{Verificatie van de correctheid van \textit{allHeapsEmpty()}}

We willen controleren of het diagram in figuur \ref{fig:nim-allHeapsEmpty} een positief antwoord geeft als alle stapels leeg zijn en een negatief antwoord als minstens \'e\'en stapel niet leeg is. We werken met een invoerstructuur die aangeeft dat de uitvoering begint met de eerste instructie van \textit{allHeapsEmpty} en eindigt na de laatste instructie van dat diagram. De invoerstructuur specificeert dat er in totaal 22 tijdstappen zijn.

We stellen vier zinnen op:

@@

De uitvoeringstijd en \textit{grounding}-grootte zijn gelijkaardig voor alle vier uitvoeringen. De uitvoering duurt ongeveer 1 minuut en 28 seconden en de \textit{grounding}-grootte is 2 249 098.

Ter vergelijking duurt het voor de theorie in bijlage \ref{app:new-nim} dat overeenkomt met het diagram in figuur \ref{fig:new-nim} ongeveer 31,6 seconden om alle modellen te vinden voor een volledig spelverloop met twee stapels, waarbij \'e\'en stapel twee objecten heeft en de andere drie. De \textit{grounding}-grootte is daarbij 142 510. De invoerstructuur voor deze uitvoering specifieert dat er in totaal 100 tijdstappen zijn.

@@

Het is mogelijk om voor een diagram bepaalde gewenste eigenschappen te controleren. Dit gaat echter gepaard met grote \textit{groundings} en een lange uitvoeringstijd.
